References for PloS paper

### 2014-02-27 image_analysis_paper/background_reading/article_notes.tex
Updated paper again.

\section{References from Stas}

@@

@@

\item \fullcite{chen2012nonlinear}

\item \fullcite{das2013dynamic}

\item \fullcite{schindler2007inferring}
\begin{itemize}
	\item Want to order images that span many years
	\item Solve a constraint satisfaction problem (CSP)
	\item Find features in images, and then try to match up the features between the different images
	\item Use structure from motion (SFM) algorithms to construct 3D points and camera angles from 2D images
	\item Given the sets of features in each image, they then try to order the images in a way that is ``most consistent'' with how buildings are constructed in the real world (i.e. buildings are built, persist for some time, and then can be torn down)-- this is what they call the CSP
\end{itemize}
\item \fullcite{bigot2012geometric}

\item \fullcite{escudero2011epithelial}

\item \fullcite{machta2013parameter}

@@

@@

\item \fullcite{xu2012comprehensive}
\begin{itemize}
	\item Want to quantify shape of liver lesions
	\item Take set of 14 common shape descriptors
	\item Use regression to find optimal combination of shape descriptors that can predict the similairity using a ``similairity reference standard''
	\item FINISH
\end{itemize}
\item \fullcite{sisan2010event}
\begin{itemize}
	\item Looks at time correlations in data to determine temporal ordering
	\item Goal is to determine in what order cellular events happen (i.e. does rxn A lead to rxn B, or does B lead to A)
	\item They look at the time-correlation between events.
	\item Asymmetry in time-correlation tells you in what order events happen (depending on model)
	\item Looking at time correlation over many samples integrates/averages out stocahsticity in system
\end{itemize}

\item \fullcite{golparvar2011monitoring}

\item \fullcite{golani1999phenotyping}
\begin{itemize}
	\item Want to construct video of construction progress from snapshots
	\item Building Information Models (BIM) 
\end{itemize}

\item \fullcite{anavy2014blind}

\item \fullcite{basu2014detecting}

@@

\section{Algorithms}

@@

\section{Applications}

>>>>

\section{Biology Applications}

@@

%\item \fullcite{golani1999phenotyping}

\item \fullcite{anavy2014blind}
\begin{itemize}
	\item Want to temporally order RNA-Seq data
	\item First do PCA on RNA-Seq data to denoise/reduce dimension
	\item Then solve a TSP in this reduced space; assume that the path found by the TSP is the ``correct'' developmental path
	\item Assume later stages in development are characterized by more variability/higher entropy to determine where the path starts (in time) and where it ends
\end{itemize}

\item \fullcite{basu2014detecting}
\begin{itemize}
	\item Want to detect changes in cell morphology
	\item First use image segmentation to extract individual cells from image
	\item Center images, and align main axis of cell to be vertical
	\item Consider pictures as mass distributions (more color = more mass); describe distance between cells to be the amount of effort one must exert (mass * distance) to make one picture look like the other
	\item Then use PCA to visualize this data
\end{itemize}

\item \fullcite{escudero2011epithelial}
\begin{itemize}
	\item Want to be able to characterize differences in cellular organization between organisms
	\item Describe cells using features of cell morphology (i.e. average size of cells) as well as features of cell organization (i.e. how many neighbors does each cell have?)
	\item Think of cell organization as a network (each cell is a node, connected by links to its neighbors), and then use network descriptors (degree, average degree of neighbors, clustering coefficient) to describe cell organization
	\item Then use PCA and discriminant analysis (like regression, but for categorize variables) to show they can separate mutants and WT cells using these network + morphology features
\end{itemize}

@@

\item \fullcite{sisan2010event}
\begin{itemize}
	\item Looks at time correlations in data to determine temporal ordering
	\item Goal is to determine in what order cellular events happen (i.e. does rxn A lead to rxn B, or does B lead to A)
	\item They look at the time-correlation between events.
	\item Asymmetry in time-correlation tells you in what order events happen (depending on model)
	\item Looking at time correlation over many samples integrates/averages out stocahsticity in system
\end{itemize}

@@

\section{Medical Statistics Applications}

\item \fullcite{das2013dynamic}
\begin{itemize}
	\item Use spline regression to model disease progression from irregular longitudinal data
\end{itemize}

\item \fullcite{xu2012comprehensive}
\begin{itemize}
	\item Want to quantify shape of liver lesions
	\item Take set of 14 common shape descriptors
	\item Use regression to find optimal combination of shape descriptors that can predict the similairity using a ``similairity reference standard''
\end{itemize}

\section{Image Applications}

\item \fullcite{schindler2007inferring}
\begin{itemize}
	\item Want to order images that span many years
	\item Solve a constraint satisfaction problem (CSP)
	\item Find features in images, and then try to match up the features between the different images
	\item Use structure from motion (SFM) algorithms to construct 3D points and camera angles from 2D images
	\item Given the sets of features in each image, they then try to order the images in a way that is ``most consistent'' with how buildings are constructed in the real world (i.e. buildings are built, persist for some time, and then can be torn down)-- this is what they call the CSP
\end{itemize}

\item \fullcite{bigot2012geometric}
\begin{itemize}
	\item Want to do PCA on 2D images (e.g. handwritten digits)
	\item Typically, Euclidean distance is not a good metric for images
	\item They define a {\em deformation operator}, and compute the deformation operator needed to take each image to a specified template
	\item They then do PCA on these deformation operators to see what are ``typical'' deformations in handwritten digits
\end{itemize}

\item \fullcite{golparvar2011monitoring}
\begin{itemize}
	\item Want to construct video of construction progress from snapshots
	\item Building Information Models (BIM) 
\end{itemize}

\section{Math/Algorithms (Temporal)}

\item \fullcite{machta2013parameter}
\begin{itemize}
	\item Look at Fisher Information Matrix (measures sensitvity of model to changes in parameters)
	\item Eigenvalues of FIM tell you how many ``important'' or stiff parameters there are in the system
	\item Demonstrate how we can use this to see how many parameters are necessary to describe diffusion and an Ising model
\end{itemize}

@@

\item \fullcite{chen2012nonlinear}
\begin{itemize}
	\item Want to look at reduced representations for data, where each data point is a function (e.g. a growth trajectory or curve)
\end{itemize}

@@

\section{Math/Algorithms (Symmetries)}

@@

\section{Applications (Symmetries)}

### 2014-02-27 image_analysis_paper/math_description.tex
Updated paper again.

@@

@@

\subsection{Alignment Using Angular Synchronization}

Often when one collects images in experiments, each image is taken in a different orientation. 
%
Therefore, the first task in image analysis is to align the set of images. 
%
We will consider settings when alignment is done with respect to two--dimensional translations and rotations of the images, as each ``object of interest'' (i.e. each embryo) is in a different position and rotated at a different angle within the image frame.

There are many ways to align images. 
%
``Template--based'' alignment \cite{ahuja2007template} is one technique, where one selects a specific image or {\em template}, and then tries to optimally align each image to that template.
%
However, this method can result in many misalignemnts for noisy images, and can be problematic if a ``good'' template image is not known {\em a priori}. 
%
Instead, we will use a technique called angular synchronization\cite{singer2011angular} to align sets of images.
%
Angular synchronization first computes the pairwise alignments for every possible pair of images, effectively using each image as a template for every other image.
%
From this large collection of pairwise alignments, it then tries to find optimal alignment for each image, such that these individual alignments are most consistent (in some metric) with the pairwise alignment.
%
Because we consider {\em all} pairwise alignments, angular synchronization is often much more robust to noise.

We will first discuss angular synchronization with {\em only} rotations. 
%
Let $I_1, \dots, I_m \in \mathbb{R}^{n \times n}$ denote the (grayscale) images that we wish to align with respect to rotations.
%
We assume that each image $I_i$ is a noisy rotated copy of the underlying image $I_{true}$ (which we are {\em not} given), such that 
\begin{equation}
I_i = R_i I_{true} + \xi_i,
\end{equation}
where $R_i \in \mathbb{R}^{2 \times 2}$ is an unknown rotation matrix and $\xi_i$ is random (often Gaussian) noise \footnote{The rotation matrices $R_i$ do not operate directly on the $n \times n$ image matrix $I_{true}$; however, we feel that this abuse of notation makes our description of the problem more intuitive.}. 
%
Our goal is to obtain estimates of $R_1, \dots, R_m$ so that we can rotate the images $I_1, \dots, I_m$ to be aligned.
%
Up to noise, 
\begin{equation} \label{eq:pairwise_rot}
I_i \approx R_i R_j^T I_j;
\end{equation}
 note that \eqref{eq:pairwise_rot} does not require knowledge of the true underlying image $I_{true}$.
%
We can obtain {\em estimates} of $R_i R_j^T$ by computing the rotations to optimally align $I_j$ to $I_i$. 
%
Let $R_{ij}$ denote the rotation that optimally aligns $I_j$ to $I_i$.
%
We then construct the matrix $H \in \mathbb{R}^{2m \times 2m}$, where $H$ is an $m \times m$ matrix of $2 \times 2$ blocks with
\begin{equation} \label{eq:H_to_R}
H_{ij} = R_{ij}
\end{equation}
%
We note that, under our assumptions, $H_{ij} \approx R_i R_j^T$, and so
\begin{equation} \label{eq:H_low_rank}
	H \approx 
	\begin{bmatrix}
	R_1 \\
	R_2 \\
	\vdots \\
	R_m
	\end{bmatrix}
	\begin{bmatrix}
	R_1^T R_2^T \dots R_m^T
	\end{bmatrix}
\end{equation}
%
We can (approximately) recover $R_1, R_2, \dots, R_m$ by computing the top block eigenvector of $H$ (i.e the eigenvectors corresponding to the two largest eigenvalues, stacked into a $2m \times 2$ matrix), and register the images by rotating image $i$ by $R_i^T$. 
%
Furthermore, this eigendecomposition also considers {\em higher--order} consistency information. 
%
For example, given our pairwise estimates $R_{ij}$, we know that relationships of the form
\begin{equation} \label{eq:triplet_consistency}
R_{ik} R_{kj} \approx R_i R_k^T R_k R_j^T = R_i R_j^T
\end{equation}
should also hold.
%
However, 
\begin{equation}
(H^2)_{ij} = \sum_k R_{ik} R_{kj}
\end{equation}
and so {\em all} infomation of the form in \eqref{eq:triplet_consistency} is contained in the matrix $H^2$.
%
Because $H$ and $H^2$ have the same eigenvectors, our problem formulation accounts for not only pairwise alignment information, but also these higher--order considerations.

We now note that, for \eqref{eq:H_low_rank} to be satisfied, we require that the group operation matrices be orthogonal, i.e. $R_i^{-1} = R_i^T$. 
%
The typical matrix representation for $ISO(2)$, the group of two--dimensional translations and rotations (using the matrix representation for the affine group \cite{...}), does not satisfy this property. 
%
However, we can (approximately) represent rotations and translations in two--dimensions using three--dimensional rotation matrices.
%
We do this by projecting the image onto a portion of the surface of a (three--dimensional) sphere.
%
Rotations in two--dimensions now correspond to rotations around one principal axis in three--dimensions, and translations in two--dimensions correspond (approximately) to rotations around the other two principal axes in three--dimensions \footnote{Clearly, not all translations can be described this way, as translations can range from $-\infty$ to $+ \infty$, and rotations are only from $0$ to $2 \pi$. However, the images we will be considering are already mostly centered, and so we are only interested in small translations that are well within the $[0, 2\pi)$ range.}.
%
These rotation matrices are (by definition) orthoganol, and successive applications of various translations and rotations can be described via multiplication of the corresponding rotation matrices in $SO(3)$.

The first step in angular synchronization is to compute the translations and rotations to optimally align each pair of data points. 
%
For each image pair $I_i$ and $I_j$, we solve
\begin{equation}\label{eq:opt_pairwise}
(\theta_{ij}, dx_{ij}, dy_{ij}) = \argmin_{
\begin{matrix}
\theta \in [0, 2\pi) \\ 
dx \in [-\Delta, \Delta]\\ 
dy \in [-\Delta, \Delta]
\end{matrix}
} \|f(I_j, \theta, dx, dy) - I_i \|_F^2.
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm, $\Delta$ is the maximum number of pixels by which we are allowed to translate the image, and $f(I_j, \theta, dx, dy)$ is a function that first rotates the image $I_j$ by $\theta$ degrees, then translates the image by $dx$ pixels in the $x$ direction, and finally translates the image by $dy$ pixels in the $y$ direction. 
%
We then compute the three--dimensional rotation, $R_{ij} \in SO(3)$, from the two--dimensional translation and rotation parameters $\theta_{ij}$, $dx_{ij}$, and $dy_{ij}$ (see Appendix for more details).
%
We then build the matrix $H \in \mathbb{R}^{3m \times 3m}$ as described in \eqref{eq:H_to_R}, and compute the top block eigenvector of $H$, $\hat{R} \in \mathbb{R}^{3m \times 3}.$ 
%
The $3 \times 3$ blocks of $\hat{R}$ contain approximations of $R_1, \dots, R_m$.  
%
Because the blocks are not guaranteed to lie in $SO(3)$, we then project each block back to $SO(3)$ (see Appendix).
%
To register the images, we compute the translations and rotation corresponding to $R_i^T$.

@@

Given our set of aligned images, we now wish to order them {\em in time} so that we can reconstruct the developmental dynamics.
%
We will use diffusion maps (DMAPS) \cite{coifman2005geometric} to temporally order our data.
%
DMAPS aims to uncover a parameterization of high-dimensional data sampled from a low-dimensional nonlinear manifold.
%
The {\em essential} requirement for DMAPS is an appropriate distance metric $d(I_i, I_j)$ for comparing data points. 
%
This can as simple as the standard Euclidean distance, or a more complex metric (such as a distance between features of the data points) for other data sets.

@@

@@

Many times, we would to both align and order our data.
%
Our proposed alignment methodology, angular synchronization, utilizes the eigendecompostion of the matrix $H \in \mathbb{R}^{3m \times 3m}$, and our proposed ordering algorithm, diffusion maps, utilizes the eigendecomposition of the matrix $A \in \mathbb{R}^{m \times m}$.
%
We can combine the two steps into one eigencomputation that allows us to {\em simultaneously} recover the optimal alignments and the temporal ordering.
%
This technique is called vector diffusion maps \cite{singer2012vector}.
%
It not only reduces the number of required eigencomputations, but also is much more robust to noise, as mistakes in alignment are also considered when ordering the data.

For vector diffusion maps, we construct the matrix $S \in \mathbb{R}^{3m \times 3m}$, with
\begin{equation}
	S_{ij} = A_{ij} H_{ij}
\end{equation}
%
where $A$ is defined in \eqref{eq:dmaps_A}, the distance metric $d(I_i, I_j)$ is the Frobenius norm between the images {\em after} optimal pairwise alignment ($d(I_i, I_j) = \| f(I_j, \theta_{ij}, dx_{ij}, dy_{ij}) - I_i \|_F$), and $H$ is defined in \eqref{eq:H_to_R}.

We then compute the eigenvalues $\lambda_1, \lambda_2, \dots, \lambda_{3m}$ and eigenvectors $\phi_1, \phi_2, \dots, \phi_{3m}$ of $S$, and order them such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_{3m}|$.
%
Again, the top (block) eigenvector of $S$ gives us approximations to the optimal rotations for each image.
%
However, we now also obtain embedding coordinates for our images.
%
In general, the embedding coordinates are given by 
\begin{equation}
\psi_{k,l} (i) = \langle \phi_k(i), \phi_l(i) \rangle
\end{equation}
where $\phi_k(i) \in \mathbb{R}^3$ denotes the $i^{th}$ block of $\phi_k$.
%
If we assume that our data are one--dimensional, and that the rotations and the dynamics are uncoupled and therefore separable, one can show that the ``correct'' embedding coordinate (i.e. the analog of $\phi_2$ from the diffusion maps case) is $\psi_{1,4}$.
%
If no such claims can be made, then one often runs a second step of diffusion maps, using $\psi_{k,l}$ as the data points, in order to uncover the true manifold modulo symmetries.

\section{Appendix}

\subsection{Angular Synchronization Implementation}

\subsubsection{Computing Optimal Pairwise Alignments}

@@

@@

>>>>

\usepackage{graphicx}
\usepackage{subcaption}
\graphicspath{ {paper_figures/} }

@@

\begin{figure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{fig2a}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{fig2b}
\caption{}
\end{subfigure}
\caption{}
\end{figure}

@@

\subsection{Registration Using Angular Synchronization} \label{subsec:angsynch}

Often when one collects images in experiments, each image is taken in a different orientation. 
%
Therefore, the first task in image analysis is to align or {\em register} the set of images. 
%
There are many ways to align images. 
%
``Template--based'' alignment \cite{ahuja2007template} is one technique, where one selects a specific image or {\em template}, and then tries to optimally align each image to that template.
%
However, this method can result in many misalignemnts for noisy images, and can be problematic if a ``good'' template image is not known {\em a priori}. 
%
Instead, we will use a technique called angular synchronization\cite{singer2011angular} to align sets of images.
%
Angular synchronization first computes the pairwise alignments for every possible pair of images, effectively using each image as a template for every other image.
%
From this large collection of pairwise alignments, it then tries to find optimal alignment for each image, such that these individual alignments are most consistent (in some metric) with the pairwise alignment.
%
Because we consider {\em all} pairwise alignments, angular synchronization is often much more robust to noise.

We will first discuss angular synchronization with {\em only} rotations. 
%
Let $x_1, \dots, x_m \in \mathbb{R}^x$ denote the (grayscale) signal that we wish to align with respect to rotations.
%
We assume that each signal $x_i$ is a noisy rotated copy of the underlying signal $x_{true}$ (which we are {\em not} given), such that 
\begin{equation}
x_i = f(x_{true}, \theta_i) + \xi_i
\end{equation}
where $f(x_{true}, \theta_i)$ rotates the signal $x_{true}$ by $\theta_i$ degrees \footnote{add about how we rotate, because it's soooooo complicated}. 
%
Our goal is to obtain estimates of $\theta_1, \dots, \theta_m$ so that we can align $x_1, \dots, x_m$. 
%
Up to noise, 
\begin{equation} \label{eq:pairwise_rot}
x_i \approx f(x_j, \theta_i - \theta_j) ;
\end{equation}
 note that \eqref{eq:pairwise_rot} does not require knowledge of $x_{true}$.
%
We can obtain {\em estimates} of $\theta_i - \theta_j$ by computing the rotations to optimally align $x_j$ to $x_i$. 
%
Let $\theta_{ij}$ denote the rotation that optimally aligns $x_j$ to $x_i$, i.e.,
%
\begin{equation} \label{eq:opt_angle}
\theta_{ij} = \argmin_{\theta} \|x_i - f(x_j, \theta)\|^2
\end{equation}

Rather than work with the angles directly, it will be more convenient to work with the rotation matrices 
\begin{equation} \label{eq:R_theta}
R(\theta_{ij}) = \begin{bmatrix}
\cos(\theta_{ij}) & \sin(\theta_{ij}) \\
-\sin(\theta_{ij}) & \cos(\theta_{ij})
\end{bmatrix}
\end{equation}
%
Successive rotations correspond to multiplication of the rotation matrices, i.e., 
$R(\alpha_1 + \alpha_2) = R(\alpha_1) R(\alpha_2)$,
and, due to the orthogonality of rotation matrices, $R(-\alpha) = R(\alpha)^T$.

Let $d$ denote the dimension of the rotation matrices we are considering (for our example of planar rotations, $d=2$).
%
 We then construct the matrix $H \in \mathbb{R}^{md \times md}$, where $H$ is an $m \times m$ matrix of $d \times d$ blocks with
\begin{equation} \label{eq:H_to_R}
H_{ij} = R(\theta_{ij})
\end{equation}
%

We note that, under our assumptions, 
\begin{equation} 
H_{ij} = R(\theta_{ij}) \approx R(\theta_i - \theta_j) = R(\theta_i) R(-\theta_j) = R(\theta_i) R(\theta_j)^T,
\end{equation}
 and so
\begin{equation} \label{eq:H_low_rank}
	H \approx 
	\begin{bmatrix}
	R(\theta_1) \\
	R(\theta_2) \\
	\vdots \\
	R(\theta_m)
	\end{bmatrix}
	\begin{bmatrix}
	R(\theta_1)^T R(\theta_2)^T \dots R(\theta_m)^T
	\end{bmatrix}
\end{equation}
%
We can (approximately) recover $R(\theta_1), R(\theta_2), \dots, R(\theta_m)$ by computing the top block eigenvector of $H$ (i.e the eigenvectors corresponding to the $d$ largest eigenvalues, stacked into a $md \times 2$ matrix).
%
We can estimate $\theta_{i}$ by inverting \eqref{eq:R_theta}, and then register the signals by rotating image $i$ by $-\theta_i$.

Furthermore, this eigendecomposition also considers {\em higher--order} consistency information. 
%
For example, given our pairwise estimates $R_{ij}$, we know that relationships of the form
\begin{equation} \label{eq:triplet_consistency}
R(\theta_{ik}) R(\theta_{kj}) \approx R(\theta_i) R(\theta_k)^T R(\theta_k) R(\theta_j)^T = R(\theta_i) R(\theta_j)^T
\end{equation}
should also hold.
%
However, 
\begin{equation}
(H^2)_{ij} = \sum_k R(\theta_{ik}) R(\theta_{kj})
\end{equation}
and so {\em all} infomation of the form in \eqref{eq:triplet_consistency} is contained in the matrix $H^2$.
%
Because $H$ and $H^2$ have the same eigenvectors, our problem formulation accounts for not only pairwise alignment information, but also these higher--order considerations.

To illustrate how angular synchronization works, we have converted the two--dimensional images into concentration profiles on a ring (we will address the full two--dimensional images in a subsequent section). 
%
Figure \ref{subfig:1d_image_example} shows a typical image that we have converted to a one--dimensional concentration profile. 
%
The images are not aligned, and so rotation of this image corresponds to shifting (with periodic boundary conditions) the one--dimensional concentration profile. 
%
Figure \ref{subfig:1d_unaligned_unordered} shows the concentration profiles for all the images in our data set. 
%
The concentration profiles have been stacked in an array, such that each row corresponds to one image.
%
Figure \ref{subfig:1d_aligned_unordered} shows the concentration profiles, now aligned using angular synchronization. 
%
The peaks are now (mostly) aligned.

\begin{figure}
\begin{subfigure}{0.17\textwidth}
\includegraphics[width=\textwidth]{illustrate_1d}
\caption{}
\label{subfig:1d_image_example}
\end{subfigure}
\begin{subfigure}{0.2\textwidth}
\includegraphics[width=\textwidth]{unregistered_unordered_1d}
\caption{}
\label{subfig:1d_unaligned_unordered}
\end{subfigure}
\begin{subfigure}{0.2\textwidth}
\includegraphics[width=\textwidth]{registered_unordered_1d}
\caption{}
\label{subfig:1d_aligned_unordered}
\end{subfigure}
\begin{subfigure}{0.2\textwidth}
\includegraphics[width=\textwidth]{registered_ordered_1d}
\caption{}
\label{subfig:1d_aligned_ordered}
\end{subfigure}
\begin{subfigure}{0.2\textwidth}
\includegraphics[width=\textwidth]{registered_ordered_vdm_1d}
\caption{}
\label{subfig:1d_aligned_ordered_vdm}
\end{subfigure}
\caption{}
\label{fig:1d_demo}
\end{figure}

@@

Given our set of registered signals, shown in Figure \ref{subfig:1d_aligned_unordered}, we now wish to order them {\em in time} so that we can reconstruct the developmental dynamics.
%
We will use diffusion maps (DMAPS) \cite{coifman2005geometric} to temporally order our data.
%
DMAPS aims to uncover a parameterization of high-dimensional data sampled from a low-dimensional nonlinear manifold.
%
The {\em essential} requirement for DMAPS is an appropriate distance metric $d(x_i, x_j)$ for comparing data points. 
%
This can as simple as the standard Euclidean distance, or a more complex metric (such as a distance between features of the data points) for other data sets.

@@

We demonstrate how we can do this with our one--dimensional concentration profiles. 
%
We use the Euclidean distance between the aligned concentration profiles (the rows of Figure \ref{subfig:1d_aligned_unordered}) as our distances in \eqref{eq:dmaps_W}.
%
Figure \ref{subfig:1d_aligned_ordered} shows the data from Figure \ref{subfig:1d_aligned_unordered}, now sorted by the value of the first (non--trivial) dmaps component, $\phi_2$.

@@

Many times, we would to both align and order our data.
%
Our proposed alignment methodology, angular synchronization, utilizes the eigendecompostion of the matrix $H \in \mathbb{R}^{md \times md}$, and our proposed ordering algorithm, diffusion maps, utilizes the eigendecomposition of the matrix $A \in \mathbb{R}^{m \times m}$.
%
We can combine the two steps into one eigencomputation that allows us to {\em simultaneously} recover the optimal alignments and the temporal ordering.
%
This technique is called vector diffusion maps (VDM) \cite{singer2012vector}.
%
It not only reduces the number of required eigencomputations, but also is much more robust to noise, as mistakes in alignment are also considered when ordering the data.

For vector diffusion maps, we construct the matrix $S \in \mathbb{R}^{md \times md}$, with
\begin{equation}
	S_{ij} = A_{ij} H_{ij}
\end{equation}
%
where $A$ is defined in \eqref{eq:dmaps_A} and $H$ is defined in \eqref{eq:H_to_R}.
%
The distance $d(x_i, x_j)$ is the distance between samples {\em after} alignment (i.e. it is the minimium obtained in \eqref{eq:opt_pairwise}).

We then compute the eigenvalues $\lambda_1, \lambda_2, \dots, \lambda_{md}$ and eigenvectors $\phi_1, \phi_2, \dots, \phi_{md}$ of $S$, and order them such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_{md}|$.
%
Again, the top (block) eigenvector of $S$ gives us approximations to the optimal rotations for each image.
%
However, we now also obtain embedding coordinates for our images.
%
In general, the embedding coordinates are given by 
\begin{equation}
\psi_{k,l} (i) = \langle \phi_k(i), \phi_l(i) \rangle
\end{equation}
where $\phi_k(i) \in \mathbb{R}^d$ denotes the $i^{th}$ block of $\phi_k$.
%
If we assume that our data are one--dimensional, and that the rotations and the dynamics are uncoupled and therefore separable, one can show that the ``correct'' embedding coordinate (i.e. the analog of $\phi_2$ from the diffusion maps case) is $\psi_{1,d+1}$.
%
If no such claims can be made, then one often runs a second step of diffusion maps, using $\psi_{k,l}$ as the data points, in order to uncover the true manifold modulo symmetries.

Figure \ref{subfig:1d_aligned_ordered_vdm} shows the result of aligning and ordering the data in Figure \ref{subfig:1d_unaligned_unordered} using vector diffusion maps. 
%
The data has been ordered by $\psi_{1, 3}$.

\subsection{Registering Images with Respect to Translations and Rotations}

We now note that, for \eqref{eq:H_low_rank} to be satisfied, we require that the group operation matrices be orthogonal, i.e. $R(\alpha)^{-1} = R(\alpha)^T$. 
%
The typical matrix representation for $ISO(2)$, the group of two--dimensional translations and rotations (using the matrix representation for the affine group \cite{...}), does not satisfy this property. 
%
However, we can (approximately) represent rotations and translations in two--dimensions using three--dimensional rotation matrices.
%
We do this by projecting the image onto a portion of the surface of a (three--dimensional) sphere.
%
Rotations in two--dimensions now correspond to rotations around one principal axis in three--dimensions, and translations in two--dimensions correspond (approximately) to rotations around the other two principal axes in three--dimensions \footnote{Clearly, not all translations can be described this way, as translations can range from $-\infty$ to $+ \infty$, and rotations are only from $0$ to $2 \pi$. However, the images we will be considering are already mostly centered, and so we are only interested in small translations that are well within the $[0, 2\pi)$ range.}.
%
These rotation matrices are (by definition) orthoganol, and successive applications of various translations and rotations can be described via multiplication of the corresponding rotation matrices in $SO(3)$.
%
See the Appendix for more details

\section{Results}

\begin{figure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{fig4a}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{dpERK_rank_corr}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{fig4c}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{Dl_rank_corr}
\caption{}
\end{subfigure}
\end{figure}

\appendix

\section{Computing Optimal Alignments for One--Dimensional Signals}

To compute the optimal angles in \eqref{eq:opt_angle}, we search over all possible shifts of the signals.

\subsubsection{Computing Optimal Pairwise Alignments}
The first step in angular synchronization is to compute the translations and rotations to optimally align each pair of data points. 
%
For each image pair $I_i$ and $I_j$, we solve
\begin{equation}\label{eq:opt_pairwise}
(\theta_{ij}, dx_{ij}, dy_{ij}) = \argmin_{
\begin{matrix}
\theta \in [0, 2\pi) \\ 
dx \in [-\Delta, \Delta]\\ 
dy \in [-\Delta, \Delta]
\end{matrix}
} \|f(I_j, \theta, dx, dy) - I_i \|_F^2.
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm, $\Delta$ is the maximum number of pixels by which we are allowed to translate the image, and $f(I_j, \theta, dx, dy)$ is a function that first rotates the image $I_j$ by $\theta$ degrees, then translates the image by $dx$ pixels in the $x$ direction, and finally translates the image by $dy$ pixels in the $y$ direction. 
%

@@

ADD HOW WE ROTATE AND TRANSLATE, FILL IN MISSING PIXELS, ETC.

@@

%
We then compute the three--dimensional rotation, $R_{ij} \in SO(3)$, from the two--dimensional translation and rotation parameters $\theta_{ij}$, $dx_{ij}$, and $dy_{ij}$ (see Appendix for more details).
%
We then build the matrix $H \in \mathbb{R}^{3m \times 3m}$ as described in \eqref{eq:H_to_R}, and compute the top block eigenvector of $H$, $\hat{R} \in \mathbb{R}^{3m \times 3}.$ 
%
The $3 \times 3$ blocks of $\hat{R}$ contain approximations of $R_1, \dots, R_m$.  
%
Because the blocks are not guaranteed to lie in $SO(3)$, we then project each block back to $SO(3)$ (see Appendix).
%
To register the images, we compute the translations and rotation corresponding to $R_i^T$.
